BIP Architectures and Open pNets

### 2019-01-29 2018-Arch-pNets/arch-pNets.tex

\addSB{Although we define properties as state predicates, any
  appropriate logic can be used to specify them.  For instance, the
  property \emph{``There is always a possibility to reset the system
    or resume operation after a single failure'' (\ie without
    additional failures occuring in the meantime)} enforced by the
  Failure Monitor architecture can be specified using CTL as $\AG
  \bigl(\PortFail \rightarrow \EU[]{\lnot \PortFail}{(\PortResume \lor
    \PortReset)}\bigr)$.  Additional properties are provided in
  \app{properties}.}

@@

\begin{example}[An open-transition]
  \label{OT:enable-composed}
  The \texttt{Failure TimerMonitor} pNet of \Fig{FailureTimer:pNet} has \replaceSB{2 controllers 
  and 2 holes. O}{two coordinators and four dangling ports, which can be grouped into two holes (see \secn{encoding} below).  Disregarding the maximal progress priority, o}ne of its possible open-transitions is:\\[-2ex]
%%   \footnote{Note the underlined action notation that we use to represent synchronised actions, i.e. actions that are the result of a synchronisation and that we declare as \emph{non-observable}. Those actions are only used for monitoring purposes and statement of properties but cannot be synchronised again in a synchronisation vector. This allows us to define a weak bisimilarity equivalence with p=compositional properties.}
%%   : 
%% \noteLH{underline notation might be a bit too much for this paper : see at the end if we remove it or not (will we use it?)}
% \smallskip
%$  OT_1  = \openrule
%{\emptyset, \{t:=0, z:=TopInt, zone:=[minZone,maxZone]\} }
%{\ostate{00} \OTarrow{\underline{\alpha_4}} \ostate{11}}
%$
%\medskip
\begin{small}\[ \openrule
  {\{B\mapsto fail(b0)\},\\
    t<GetMax(z) \land b1=true \land b1=>b0 \land z'= (z \cap (b1?zone;TopInt)+t) {\color{blue}\land (t' = t) \land (zone' = zone)},\\
    \{z:=z', {\color{blue}t := t'+1, zone := zone'}\} }
  {\ostate{11} \OTarrow{\alpha_0} \ostate{12}}
  \]
\end{small}\vspace{-1.5ex}
%
%The first one, $OT_1$, corresponds to the simultaneous initialisation
%of both controlers. No hole is active at that time, and the resulting
%global action is denoted $\alpha_4$.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\columnwidth]{FailureTimerPNET-v4}
  \caption{The open pNET encoding the Failure Monitor architecture (\fig{schema:ArchFailure:BIP}) without the priority model}
  \label{fig:FailureTimer:pNet}
%
  \noteSBin{The updated encoding exports more primed variables in the
    actions, \eg $tick(z,t,z',t')$ and $fail(zone, zone', b1)$ instead
    of $tick(z,t,z')$ and $fail(zone, b1)$, and SVs, \eg $\dots \land
    (t' = t) \land (zone' = zone)$ in the guard of SV0. Shall we
    update the figure to include these or, rather, explain that the
    figure is cleaned-out for readability?}
%
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\columnwidth]{TimerOADetailed}
  \caption{The open Automaton of the Failure Monitor Architecture}
  \label{schema:ArchFailure:BIP}
\end{figure}

In Figure~\ref{schema:ArchFailure:OA} we show the full open automaton
obtained from our pNet. For reasons of space, we have not shown the
detail of the open transition, but only the assignement of state
variables, and the resulting action. You can find the full details in
Appendix \ref{secn:full-results}. As an example, $OT_3$ is the one explained in Example
\ref{Example:TimerOT3} page \pageref{Example:TimerOT3}

\todoEM{12 open transitions... Comments on the predicates ???}

>>>>

\label{property:reset}
\addSB{Although we define properties as state predicates, any
  appropriate logic can be used to specify them.  For instance, the
  property \emph{``There is always a possibility to reset the system
    or resume operation after a single failure'' (\ie without
    additional failures occuring in the meantime)} enforced by the
  Failure Monitor architecture can be specified using CTL as $\AG
  \bigl(\PortFail \rightarrow \EU[]{\lnot \PortFail}{(\PortResume \lor
    \PortReset)}\bigr)$.  Additional properties are provided in
  \app{properties}.}

@@

\begin{example}[An open-transition]
  \label{Example:TimerOT3}
  The \texttt{Failure TimerMonitor} pNet of \Fig{FailureTimer:pNet} has \replaceSB{2 controllers 
  and 2 holes. O}{two coordinators and four dangling ports, which can be grouped into two holes (see \secn{encoding} below).  Disregarding the maximal progress priority, o}ne of its possible open-transitions is:\\[-2ex]
%%   \footnote{Note the underlined action notation that we use to represent synchronised actions, i.e. actions that are the result of a synchronisation and that we declare as \emph{non-observable}. Those actions are only used for monitoring purposes and statement of properties but cannot be synchronised again in a synchronisation vector. This allows us to define a weak bisimilarity equivalence with p=compositional properties.}
%%   : 
%% \noteLH{underline notation might be a bit too much for this paper : see at the end if we remove it or not (will we use it?)}
% \smallskip
%$  OT_1  = \openrule
%{\emptyset, \{t:=0, z:=TopInt, zone:=[minZone,maxZone]\} }
%{\ostate{00} \OTarrow{\underline{\alpha_4}} \ostate{11}}
%$
%\medskip
\begin{small}\[ \openrule
  {\{B\mapsto fail(b0)\},\\
    t<GetMax(z) \land b1=true \land b1=>b0 \land z'= (z \cap (b1?zone;TopInt)+t) {\color{blue}\land (t' = t) \land (zone' = zone)},\\
    \{z:=z', {\color{blue}t := t'+1, zone := zone'}\} }
  {\ostate{11} \OTarrow{\alpha_0} \ostate{12}}
  \]
\end{small}\vspace{-1.5ex}
%
%The first one, $OT_1$, corresponds to the simultaneous initialisation
%of both controlers. No hole is active at that time, and the resulting
%global action is denoted $\alpha_4$.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\columnwidth]{FailureTimerPNET-v4}
  \caption{The open pNET encoding the Failure Monitor architecture (\fig{schema:ArchFailure:BIP}) without the priority model}
  \label{fig:FailureTimer:pNet}
%% %
%%   \noteSBin{The updated encoding exports more primed variables in the
%%     actions, \eg $tick(z,t,z',t')$ and $fail(zone, zone', b1)$ instead
%%     of $tick(z,t,z')$ and $fail(zone, b1)$, and SVs, \eg $\dots \land
%%     (t' = t) \land (zone' = zone)$ in the guard of SV0. Shall we
%%     update the figure to include these or, rather, explain that the
%%     figure is cleaned-out for readability?}
%
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\columnwidth]{TimerOADetailed}
  \caption{The open Automaton of the Failure Monitor Architecture}
  \label{schema:ArchFailure:OA}
\end{figure}

In Figure~\ref{schema:ArchFailure:OA} we show the full open automaton
obtained from our pNet. For reasons of space, we have not shown the
detail of the open transition, but only the assignement of state
variables, and the resulting action; you can find the full details in
Appendix \ref{secn:full-results}. As an example, $OT_3$ is the one
explained in Example~\ref{Example:TimerOT3} page~\pageref{Example:TimerOT3}

This automaton has 12 transitions, including those encoding various
possible firing of some interactions (e.g. OT 2 and 3 for
$fail$). At first sight, it appears the rough behaviour is
reasonable, although a detailed study shows that the property stated
in page~\pageref{property:reset} does not hold, because of the
$fail$ loops on states $11$ and $12$. This is because we did not use
the maximal progress assumption here. If we do, we get the corrected
behaviour in Fig~\ref{schema:ArchFailure:MaxProgress}. This one
verifies all the properties listed in  \app{properties}.
